|

### 2020-01-01 appendix/appendix.Rmd
add edits to appendix

In this section, we explain in detail how the _Tá de Pé_ app works. The
treatment involves the following procedures:

1. Users download the app.

2. Users access the app and add new information about school constructions
	 near to their location.

3. The app sends a message to the mayor's office requesting details about the
	 project. If necessary, the message is also forwarded to other government
	 oversight institutions such as the city council and the Brazilian Federal
	 Comptroller (CGU). Below, we discuss the cases in which we forwarded the
	 denounce to the authorities.

4. If the pressure over the mayors has an effect, we expect to witness
	 improvements on governmental data regarding the school construction
	 progress.

The treatment is a compounded treatment comprised of steps (2) to (4). To make
sure that the only source of effect comes from step (2), we took the following
precautions:

In step (1), we have implemented Facebook campaigns in municipalities that have
school construction agreements with the Federal government. We did not
discriminate any municipality in this step, and citizens living in
municipalities in the treatment and in the control could equally have received
the banners about the _Tá de Pé_ app. The Facebook campaign was not randomised.

For step (3), we put a team of two data scientists to ensure that upon
receiving a denounce, the mayor's office, the city council, and the Brazilian
Federal Comptroller (CGU) would be notified. The decision making funnel was:

3.1. We notified the _Engineers without Borders_ (EWB) about the denounce and
sent them the user-provided pictures and auxiliary information. They issued a
report on how much the construction was likely to be delayed.

3.2. After we had the EWB report, we notified the mayor's office in compliance
with Brazil's Freedom of Information Act (FOI). We communicated that a citizen
had anonymously made a denounce about a school (fully identifying the school
construction that we were referring to), and that the Engineers without Borders
found the construction delayed. We submitted a request for the mayor's office
to provide an explanation on why the construction was delayed, and asked for
updated statistics on when the construction will be finished. According to the
FOI law, the mayor had twenty days to respond.

3.3. If the response was deemed insufficient, we forwarded the denounce to
the city council and to the Brazilian Federal Comptroller (CGU). As the
denounce goes to these institutions, they can harm mayors by restricting access
to Federal funds.

3.4. At each stage, a Twitter bot constructed by Transparência Brasil
(<https://twitter.com/tadepeapp>) reported the reception of a denounce and
tracked the responses of the mayor, the city council, and the CGU.

In step (4), we relied on administrative data by the Brazilian Ministry of
Education. The Ministry of Education provides monthly data on school
constructions build with Federal Government resources and shows how many of
them are delayed across the country. We use the finishing dates reported by the
municipality at the time they sign the agreement with the Federal Government as
our measure of construction delays. This data helped us to keep the app
updated. We selected outcomes six months before and after the treatment. The
data were constantly updated by the Ministry of Education, and we found no red
flags in the overall quality of their information.

Therefore, steps (1), (3), and (4) were standardised and do not vary across the
country. The only source of variation in our intervention should come from step
(2).

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, which happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we randomly excluded selected schools out of the
app. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the app. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the app.

We argue that the app promotes two crucial features of bottom-up
accountability. First, it helps people to access information that would
otherwise be available only in an obscure spreadsheet on the Ministry of
Education website. Second, whenever people find this information troublesome,
we are providing an easy way to them to act upon the information. As the _Tá de
Pé_ design was intuitive and straightforward, we believe that most of
Brazilians would find it easy to use the app.

Moreover, the success of treatment does not require that everyone in the
municipality uses it. The more usage means more pictures and denounces, and
this represents more pressure over the mayors office, what could revert into a
higher effect. However, just one denounce would be enough for the project to be
successful, as the denounce can be directed to the Brazilian General
Comptroller (CGU), and this could potentially deny future funding to the
municipality.

@@

We collect the main outcomes using the data reported by the municipality to the
Brazilian Ministry of Education. Note that this is the same data as the
pre-treatment data that we use to feed in the app with delayed schools.

The outcomes used in the paper are:

1. **Percentage of the project completed before the impact evaluation
	 started**: This is a column in the Ministry of Education dataset. The data
	 in this column were sampled in the beginning of the intervention.

2. **Percentage of the project reported as completed by the end of the
	 intervention period**: Another column in the Ministry of Education
	 dataset. The data were sampled six months after the intervention.

3. **Difference between the percentage reported as completed before and after
	 the intervention**: The difference between (1) and (2) above. Note
	 that according to @gerber2012field, this way to look into outcomes is
	 positive, as having a pre-treatment covariate that highly predicts the
	 outcome reduces the variance in the estimator considerably.

4. **Dummy indicator for finished constructions**: The Ministry of Education
	 dataset has a column with the construction status. We created a dummy to be
	 one when the construction was reported as finished and zero otherwise. The
	 data here were sampled six months after the intervention.

5. **Dummy indicator for cancelled constructions**: Similarly, we created a
	dummy variable from the Ministry of Education dataset indicating when the
	construction was reported as cancelled. The data here were sampled six months
	after the intervention.

6. **Number of schools where construction companies updated the conclusion
	 dates**: Lastly, the Ministry of Education dataset has a column with the
	 construction estimated completion date. We took the completion dated
	 reported before and six months after the intervention. If the dates
	 differed, we coded as one, and zero otherwise.

All these variables come from our manipulation of the Ministry of Education
data. Collecting the outcomes in this way is superior than the alternative of
gathering the data ourselves. First, we would not be able to manipulate the
results and cherry-pick favourable results. Even if they have the best
intentions, researchers may bias the results just by the choices they make in
the data collection process. For instance, if we send an RA to collect the
data, as she knows the app, she could unconsciously over-report positive
results. This would bias our findings and undermine the credibility of our
experiment. Second, using external data makes the results easily reproducible
by other researchers. As the Ministry of Education has standardised data
collection procedures, they greatly simplify and facilitate the reproduction of
our findings.

@@

The app by itself would probably not have had an effect on the responsiveness
of mayors. To boost usage and publicize the app, Transparência Brasil ran 64
Facebook campaigns throughout the country. The Facebook campaign concentrates
the bulk of the investments in the beginning of the intervention. The app
download data is in the second panel of Figures 2 and 3. For the Facebook
interaction, we have the following statistics:

1. Impressions
  - Mean: `r mean(fballt$impressoes)`
  - Total for all campaigns: `r sum(fballt$impressoes)`

2. App Download link clicks:
  - Mean: `r mean(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`
  - Total for all campaigns: `r sum(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`

As we see from the actual download activity, the people who indeed downloaded
the app were much fewer than these numbers would suggest.

We have chosen Facebook because the platform offers an easy and cheap method to
spread information. Moreover, according to the Statista website, Facebook is
used by 70 million Brazilians (60% of the entire adult population), and the
demographics largely match the demographics publicised by the Brazilian
Institute of Geography and Statistics. For instance, see the following
websites:

1. [Statista Total Facebook Users](https://www.statista.com/statistics/244936/number-of-facebook-users-in-brazil/)
2. [Statista Total Facebook Users by Age](https://www.statista.com/statistics/866282/facebook-user-share-brazil-age/)

>>>>

__In this section, we explain in detail how the _Tá de Pé_ app works. The
treatment involves the following procedures:__

**1. Users download the app.**

**2. Users access the app and add new information about school constructions**
**near to their location.**

**3. The app sends a message to the mayor's office requesting details about
the project. If necessary, the message is also forwarded to other government
oversight institutions such as the city council and the Brazilian Federal
Comptroller (CGU). Below, we discuss the cases in which we forwarded the
denounce to the authorities.**

**4. If the pressure over the mayors has an effect, we expect to witness
	 improvements on governmental data regarding the school construction
	 progress.**

**The treatment is a compounded treatment comprised of steps (2) to (4). To
make sure that the only source of effect comes from step (2), we took the
following precautions:**

**In step (1), we have implemented Facebook campaigns in municipalities that
have school construction agreements with the Federal government. We did not
discriminate any municipality in this step, and citizens living in
municipalities in the treatment and in the control could equally have received
the banners about the _Tá de Pé_ app. The Facebook campaign was not
randomised.**

**For step (3), we put a team of two data scientists to ensure that upon
receiving a denounce, the mayor's office, the city council, and the Brazilian
Federal Comptroller (CGU) would be notified. The decision making funnel was:**

**3.1. We notified the _Engineers without Borders_ (EWB) about the denounce and
sent them the user-provided pictures and auxiliary information. They issued a
report on how much the construction was likely to be delayed.**

**3.2. After we had the EWB report, we notified the mayor's office in compliance
with Brazil's Freedom of Information Act (FOI). We communicated that a citizen
had anonymously made a denounce about a school (fully identifying the school
construction that we were referring to), and that the Engineers without Borders
found the construction delayed. We submitted a request for the mayor's office
to provide an explanation on why the construction was delayed, and asked for
updated statistics on when the construction will be finished. According to the
FOI law, the mayor had twenty days to respond.**

**3.3. If the response was deemed insufficient, we forwarded the denounce to
the city council and to the Brazilian Federal Comptroller (CGU). As the
denounce goes to these institutions, they can harm mayors by restricting access
to Federal funds.**

**3.4. At each stage, a Twitter bot constructed by Transparência Brasil
(<https://twitter.com/tadepeapp>) reported the reception of a denounce and
tracked the responses of the mayor, the city council, and the CGU.**

**In step (4), we relied on administrative data by the Brazilian Ministry of
Education. The Ministry of Education provides monthly data on school
constructions build with Federal Government resources and shows how many of
them are delayed across the country. We use the finishing dates reported by the
municipality at the time they sign the agreement with the Federal Government as
our measure of construction delays. This data helped us to keep the app
updated. We selected outcomes six months before and after the treatment. The
data were constantly updated by the Ministry of Education, and we found no red
flags in the overall quality of their information.**

**Therefore, steps (1), (3), and (4) were standardised and do not vary across
the country. The only source of variation in our intervention should come from
step (2).**

**In step (2), we applied the treatment by removing schools within
municipalities out of the app. In the first intervention, which happened at the
municipal level, we excluded all schools within a given randomly selected
municipality. In the second intervention, we randomly excluded selected schools
out of the app. Removing a school means that the school was in the dataset
provided by the Brazilian Ministry of Education, but it did not show up in the
app. Therefore, our treatment effect represents the school being present
(treatment) or absent (control) in the app.**

**We argue that the app promotes two crucial features of bottom-up
accountability. First, it helps people to access information that would
otherwise be available only in an obscure spreadsheet on the Ministry of
Education website. Second, whenever people find this information troublesome,
we are providing an easy way to them to act upon the information. As the _Tá de
Pé_ design was intuitive and straightforward, we believe that most of
Brazilians would find it easy to use the app.**

**Moreover, the success of treatment does not require that everyone in the
municipality uses it. The more usage means more pictures and denounces, and
this represents more pressure over the mayors office, what could revert into a
higher effect. However, just one denounce would be enough for the project to be
successful, as the denounce can be directed to the Brazilian General
Comptroller (CGU), and this could potentially deny future funding to the
municipality.**

@@

**We collect the main outcomes using the data reported by the municipality to
the Brazilian Ministry of Education. Note that this is the same data as the
pre-treatment data that we use to feed in the app with delayed schools.**

**The outcomes used in the paper are:**

**1. Percentage of the project completed before the impact evaluation started:
	 This is a column in the Ministry of Education dataset. The data in this
	 column were sampled in the beginning of the intervention.**

**2. Percentage of the project reported as completed by the end of the
	 intervention period: Another column in the Ministry of Education dataset.
	 The data were sampled six months after the intervention.**

**3. Difference between the percentage reported as completed before and after
	 the intervention: The difference between (1) and (2) above. Note
	 that according to @gerber2012field, this way to look into outcomes is
	 positive, as having a pre-treatment covariate that highly predicts the
	 outcome reduces the variance in the estimator considerably.**

**4. Dummy indicator for finished constructions: The Ministry of Education
	 dataset has a column with the construction status. We created a dummy to be
	 one when the construction was reported as finished and zero otherwise. The
	 data here were sampled six months after the intervention.**

**5. Dummy indicator for cancelled constructions: Similarly, we created a
	dummy variable from the Ministry of Education dataset indicating when the
	construction was reported as cancelled. The data here were sampled six months
	after the intervention.**

**6. Number of schools where construction companies updated the conclusion
	 dates: Lastly, the Ministry of Education dataset has a column with the
	 construction estimated completion date. We took the completion dated
	 reported before and six months after the intervention. If the dates
	 differed, we coded as one, and zero otherwise.**

**All these variables come from our manipulation of the Ministry of Education
data. Collecting the outcomes in this way is superior than the alternative of
gathering the data ourselves. First, we would not be able to manipulate the
results and cherry-pick favourable results. Even if they have the best
intentions, researchers may bias the results just by the choices they make in
the data collection process. For instance, if we send an RA to collect the
data, as she knows the app, she could unconsciously over-report positive
results. This would bias our findings and undermine the credibility of our
experiment. Second, using external data makes the results easily reproducible
by other researchers. As the Ministry of Education has standardised data
collection procedures, they greatly simplify and facilitate the reproduction of
our findings.**

@@

**The app by itself would probably not have had an effect on the responsiveness
of mayors. To boost usage and publicize the app, Transparência Brasil ran 64
Facebook campaigns throughout the country. The Facebook campaign concentrates
the bulk of the investments in the beginning of the intervention. The app
download data is in the second panel of Figures 2 and 3. For the Facebook
interaction, we have the following statistics:**

**1. Impressions**
- **Mean: `r format(round(mean(fballt$impressoes),2), big.mark = ",")`**
	- **Total for all campaigns: `r format(sum(fballt$impressoes), big.mark = ",")`**

**2. App Download link clicks:
	- Mean: `r format(mean(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular']), big.mark = ",")`**
  - **Total for all campaigns: `r format(sum(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular']), big.mark = ",")`**

**As we see from the actual download activity, the people who indeed downloaded
the app were much fewer than these numbers would suggest.**

**We have chosen Facebook because the platform offers an easy and cheap method
to spread information. Moreover, according to the Statista website, Facebook is
used by 70 million Brazilians (60% of the entire adult population), and the
demographics largely match the demographics publicised by the Brazilian
Institute of Geography and Statistics. For instance, see the following
websites:**

**1. [Statista Total Facebook
Users](https://www.statista.com/statistics/244936/number-of-facebook-users-in-brazil/)**

**2. [Statista Total Facebook Users by
	 Age](https://www.statista.com/statistics/866282/facebook-user-share-brazil-age/)**
